---
title: "Bias Busters — EEDI Mixed-Effects Analysis"
subtitle: "Statistical Modelling for Research Objectives 1–6"
output:
  html_document:
    toc: true
    toc_float: true
    code_folding: show
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE, message = FALSE, warning = FALSE)
```

## Setup

```{r packages}
# Run once if required:
# install.packages(c("readr", "dplyr", "lme4", "lmerTest", "ordinal", "emmeans", "performance", "see", "broom.mixed"))

library(readr)
library(dplyr)
library(lme4)
library(lmerTest)    # adds Satterthwaite p-values to lmer() output
library(ordinal)     # clm()/clmm() for the ordinal EEDI score
library(emmeans)
library(performance) # r2() for marginal/conditional R^2 on mixed models
library(see)         # diagnostic plots returned by performance
library(broom.mixed)
```

## Data

```{r load-data}
DATA_PATH <- "../data/judged_outputs_final.csv"

df <- read_csv(DATA_PATH, show_col_types = FALSE)

REQUIRED_COLUMNS <- c(
  "stage", "gender", "age_group", "tenure", "platform", "condition_id",
  "agency_ratio", "formality_score", "gain_loss_ratio",
  "semantic_distance", "eedi_score"
)
missing_columns <- setdiff(REQUIRED_COLUMNS, names(df))
if (length(missing_columns) > 0L) {
  stop("Missing required columns: ", paste(missing_columns, collapse = ", "))
}

# Set and validate the reference category for a variable.
relevel_checked <- function(x, reference, variable) {
  x <- factor(x)
  if (!reference %in% levels(x)) {
    stop(sprintf("Reference level '%s' is absent from %s.", reference, variable))
  }
  relevel(x, ref = reference)
}

# Match the reference condition used elsewhere in the pipeline.
df <- df %>%
  mutate(
    gender = relevel_checked(gender, "female", "gender"),
    age_group = relevel_checked(age_group, "55-64", "age_group"),
    platform = relevel_checked(platform, "gpt", "platform"),
    condition_id = factor(condition_id)
  )

stage1 <- df %>% filter(stage == "stage1")
stage2 <- df %>% filter(stage == "stage2")

stage2 <- stage2 %>%
  mutate(
    tenure = relevel_checked(
      tenure, "long-term high-value supporter", "Stage 2 tenure"
    )
  )

if (nrow(stage1) == 0L || nrow(stage2) == 0L) {
  stop("Both Stage 1 and Stage 2 must contain at least one row.")
}

METRICS <- c(
  "agency_ratio", "formality_score", "gain_loss_ratio", "semantic_distance"
)

# Candidate terms for model selection.
STAGE1_CANDIDATES <- tribble(
  ~term,
  "platform",
  "gender",
  "age_group",
  "gender:platform",
  "age_group:platform",
  "gender:age_group",
  "gender:age_group:platform"
)

STAGE2_CANDIDATES <- tribble(
  ~term,
  "platform",
  "gender",
  "age_group",
  "tenure",
  "gender:platform",
  "age_group:platform",
  "gender:age_group",
  "gender:tenure",
  "age_group:tenure",
  "platform:tenure",
  "gender:platform:tenure",
  "age_group:platform:tenure"
)

STAGE1_FULL_RHS <- paste(STAGE1_CANDIDATES$term, collapse = " + ")
STAGE2_FULL_RHS <- paste(STAGE2_CANDIDATES$term, collapse = " + ")

cat(sprintf("Stage 1: %d rows | Stage 2: %d rows\n", nrow(stage1), nrow(stage2)))
```

## Helper functions

```{r helpers}
# Print a clear heading for one linguistic metric.
metric_header <- function(m, suffix = NULL) {
  cat("\n====================", m, suffix, "====================\n")
}

# Print a heading for a set of BH-adjusted results.
bh_header <- function(label = "p-values") {
  cat(sprintf("\n==== BH-adjusted %s across all four metrics ====\n", label))
}

# Convert a drop-one table into a standard candidate-history table.
standardise_drop_table <- function(table, metric, step,
                                   removed_term = NULL,
                                   denominator_df = NULL) {
  result <- as.data.frame(table) %>%
    tibble::rownames_to_column("candidate_term") %>%
    filter(candidate_term != "<none>")
  if (nrow(result) == 0L) {
    return(tibble(
      metric = character(), step = integer(), candidate_term = character(),
      statistic = numeric(), df1 = numeric(), df2 = numeric(),
      p.value = numeric(), removed = logical()
    ))
  }
  p_column <- grep("^Pr\\(", names(result), value = TRUE)
  f_column <- intersect(c("F value", "F-value"), names(result))
  numdf_column <- intersect(c("NumDF", "Df"), names(result))
  dendf_column <- intersect(c("DenDF"), names(result))
  if (length(p_column) != 1L) {
    stop("Expected exactly one p-value column in the drop-one table.")
  }
  if (length(f_column) != 1L) {
    stop("Expected exactly one F-statistic column in the drop-one table.")
  }
  if (length(numdf_column) != 1L) {
    stop("Expected exactly one numerator degrees-of-freedom column.")
  }
  if (length(dendf_column) == 1L) {
    df2 <- result[[dendf_column]]
  } else if (!is.null(denominator_df)) {
    df2 <- rep(denominator_df, nrow(result))
  } else {
    stop("Expected denominator degrees of freedom in the drop-one table.")
  }
  if (anyNA(result[[p_column]])) {
    stop("The drop-one table contains a missing p-value.")
  }
  tibble(
    metric = metric,
    step = step,
    candidate_term = result$candidate_term,
    statistic = result[[f_column]],
    df1 = result[[numdf_column]],
    df2 = df2,
    p.value = result[[p_column]],
    removed = if (is.null(removed_term)) {
      rep(FALSE, nrow(result))
    } else {
      result$candidate_term == removed_term
    }
  )
}

# Reconstruct every mixed-model fixed-effect elimination step.
mixed_fixed_history <- function(initial_model, step_table, metric, analysis_data) {
  elimination <- as.data.frame(step_table) %>%
    tibble::rownames_to_column("candidate_term") %>%
    filter(Eliminated > 0L) %>%
    arrange(Eliminated)
  current_model <- initial_model
  history <- list()
  models <- list(step0 = current_model)
  step_number <- 1L

  for (term in elimination$candidate_term) {
    tests <- drop1(current_model, ddf = "Satterthwaite")
    history[[step_number]] <- standardise_drop_table(
      tests, metric, step_number, term
    )
    current_model <- update(
      current_model,
      as.formula(paste(". ~ . -", term)),
      data = analysis_data
    )
    models[[paste0("step", step_number)]] <- current_model
    step_number <- step_number + 1L
  }

  final_tests <- drop1(current_model, ddf = "Satterthwaite")
  history[[step_number]] <- standardise_drop_table(
    final_tests, metric, step_number
  )
  list(history = bind_rows(history), models = models, final_model = current_model)
}

# Perform backward elimination for an ordinary linear model.
backward_ols <- function(metric, full_rhs, analysis_data, alpha = 0.05) {
  current_model <- lm(
    as.formula(paste(metric, "~", full_rhs)),
    data = analysis_data
  )
  history <- list()
  models <- list(step0 = current_model)
  step_number <- 1L

  repeat {
    tests <- drop1(current_model, test = "F")
    standard <- standardise_drop_table(
      tests,
      metric,
      step_number,
      denominator_df = df.residual(current_model)
    )
    removable <- standard %>% filter(p.value > alpha)
    if (nrow(removable) == 0L) {
      history[[step_number]] <- standard
      break
    }
    removed_term <- removable$candidate_term[[which.max(removable$p.value)]]
    standard$removed[standard$candidate_term == removed_term] <- TRUE
    history[[step_number]] <- standard
    current_model <- update(
      current_model,
      as.formula(paste(". ~ . -", removed_term))
    )
    models[[paste0("step", step_number)]] <- current_model
    step_number <- step_number + 1L
  }
  list(history = bind_rows(history), models = models, final_model = current_model)
}

# Convert a random-effect test into a step-history table.
standardise_random_table <- function(table, step, removed_term = NULL) {
  as.data.frame(table) %>%
    tibble::rownames_to_column("random_term") %>%
    filter(random_term != "<none>") %>%
    mutate(
      step = step,
      removed = if (is.null(removed_term)) {
        FALSE
      } else {
        random_term == removed_term
      },
      .before = 1L
    )
}

# Extract one likelihood-ratio comparison between nested mixed models.
random_lrt_row <- function(reduced_model, full_model, removed_term) {
  comparison <- anova(reduced_model, full_model, refit = FALSE)
  p_column <- grep("^Pr\\(", names(comparison), value = TRUE)
  df_column <- intersect(c("Df", "Chi Df"), names(comparison))
  if (length(p_column) != 1L || length(df_column) != 1L) {
    stop("Unexpected likelihood-ratio comparison table.")
  }
  tibble(
    random_term = removed_term,
    npar = comparison$npar[[2L]],
    logLik = comparison$logLik[[2L]],
    AIC = comparison$AIC[[2L]],
    LRT = comparison$Chisq[[2L]],
    Df = comparison[[df_column]][[2L]],
    `Pr(>Chisq)` = comparison[[p_column]][[2L]]
  )
}

# Compare the three candidate random-effects structures using LRTs.
select_random_structure <- function(metric, full_rhs, analysis_data,
                                    alpha = 0.10) {
  random_structures <- c(
    full = "(1 | condition_id) + (1 | condition_id:platform)",
    condition = "(1 | condition_id)",
    condition_platform = "(1 | condition_id:platform)"
  )
  models <- lapply(random_structures, function(random_rhs) {
    lmer(
      as.formula(paste(metric, "~", full_rhs, "+", random_rhs)),
      data = analysis_data,
      REML = TRUE
    )
  })

  candidates <- bind_rows(
    random_lrt_row(
      models$condition,
      models$full,
      "(1 | condition_id:platform)"
    ),
    random_lrt_row(
      models$condition_platform,
      models$full,
      "(1 | condition_id)"
    )
  )
  p_column <- "Pr(>Chisq)"
  removable <- which(candidates[[p_column]] > alpha)
  if (length(removable) == 0L) {
    return(list(
      full_model = models$full,
      selected_model = models$full,
      history = candidates %>% mutate(step = 1L, removed = FALSE, .before = 1L)
    ))
  }

  remove_index <- removable[[which.max(candidates[[p_column]][removable])]]
  removed_term <- candidates$random_term[[remove_index]]
  selected_name <- if (grepl("condition_id:platform", removed_term)) {
    "condition"
  } else {
    "condition_platform"
  }
  selected_model <- models[[selected_name]]
  history <- list(
    candidates %>%
      mutate(
        step = 1L,
        removed = random_term == removed_term,
        .before = 1L
      )
  )

  remaining_test <- lmerTest::ranova(selected_model, reduce.terms = TRUE)
  remaining_candidate <- remaining_test[-1L, , drop = FALSE]
  remaining_p <- remaining_candidate[[p_column]][[1L]]
  remaining_term <- rownames(remaining_candidate)[[1L]]
  remove_remaining <- remaining_p > alpha
  history[[2L]] <- standardise_random_table(
    remaining_test,
    2L,
    if (remove_remaining) remaining_term else NULL
  )

  list(
    full_model = models$full,
    selected_model = if (remove_remaining) NULL else selected_model,
    history = bind_rows(history)
  )
}

# Simplify random effects, then fixed effects, for one metric.
backward_select_metric <- function(metric, full_rhs, analysis_data,
                                   alpha_random = 0.10,
                                   alpha_fixed = 0.05) {
  random_result <- select_random_structure(
    metric, full_rhs, analysis_data, alpha_random
  )
  full_model <- random_result$full_model
  random_model <- random_result$selected_model
  use_ols <- is.null(random_model) ||
    isSingular(random_model)

  if (use_ols) {
    fixed_result <- backward_ols(metric, full_rhs, analysis_data, alpha_fixed)
    final_model <- fixed_result$final_model
    fixed_step <- NULL
  } else {
    fixed_step <- lmerTest::step(
      random_model,
      reduce.random = FALSE,
      reduce.fixed = TRUE,
      alpha.fixed = alpha_fixed,
      ddf = "Satterthwaite"
    )
    reconstructed <- mixed_fixed_history(
      random_model,
      fixed_step$fixed,
      metric,
      analysis_data
    )
    final_model <- lmerTest::get_model(fixed_step)
    fixed_result <- reconstructed
    if (isSingular(final_model)) {
      use_ols <- TRUE
      fixed_result <- backward_ols(metric, full_rhs, analysis_data, alpha_fixed)
      final_model <- fixed_result$final_model
      fixed_step <- NULL
    }
  }

  final_terms <- attr(terms(final_model), "term.labels")
  final_rhs <- if (length(final_terms) == 0L) "1" else paste(final_terms, collapse = " + ")

  list(
    metric = metric,
    full_model = full_model,
    random_history = random_result$history,
    random_model = random_model,
    fixed_step = fixed_step,
    history = fixed_result$history,
    models = fixed_result$models,
    final_model = final_model,
    model_type = if (use_ols) "OLS" else "mixed",
    final_rhs = final_rhs
  )
}

# Run one metric-level backward-elimination job.
run_backward_metric_job <- function(metric, full_rhs, analysis_data) {
  backward_select_metric(metric, full_rhs, analysis_data)
}

# Run backward elimination in parallel across linguistic metrics.
run_backward_selection <- function(full_rhs, analysis_data, metrics) {
  detected <- parallel::detectCores(logical = FALSE)
  if (is.na(detected)) detected <- 1L
  workers <- min(length(metrics), max(1L, detected - 1L))
  cat(sprintf(
    "Running %d backward selections with %d worker(s).\n",
    length(metrics), workers
  ))
  cluster <- parallel::makeCluster(workers)
  on.exit(parallel::stopCluster(cluster), add = TRUE)
  invisible(parallel::clusterEvalQ(cluster, {
    library(dplyr)
    library(lme4)
    library(lmerTest)
  }))
  parallel::clusterExport(
    cluster,
    c(
      "standardise_drop_table",
      "mixed_fixed_history", "standardise_random_table",
      "random_lrt_row", "select_random_structure",
      "backward_ols", "backward_select_metric", "run_backward_metric_job"
    ),
    envir = environment()
  )
  results <- parallel::parLapply(
    cluster, metrics, run_backward_metric_job, full_rhs, analysis_data
  )
  setNames(results, metrics)
}

# Combine complete fixed-effect elimination histories.
backward_history_table <- function(selection_results) {
  bind_rows(lapply(selection_results, `[[`, "history")) %>%
    arrange(metric, step, desc(p.value))
}

# Summarise the term removed at every successful step.
backward_path_table <- function(selection_results) {
  backward_history_table(selection_results) %>% filter(removed)
}

# Extract final fitted models.
backward_final_models <- function(selection_results) {
  setNames(lapply(selection_results, `[[`, "final_model"), names(selection_results))
}

# Combine random-effect elimination histories across metrics.
random_elimination_table <- function(selection_results) {
  bind_rows(lapply(selection_results, function(result) {
    result$random_history %>%
      mutate(metric = result$metric, .before = 1L)
  }))
}

# Summarise the change in explained variance after each term is removed.
backward_contribution_table <- function(selection_results) {
  bind_rows(lapply(selection_results, function(result) {
    path <- result$history %>% filter(removed) %>% arrange(step)
    if (nrow(path) == 0L) return(tibble())
    bind_rows(lapply(seq_len(nrow(path)), function(i) {
      before <- model_r2(result$models[[i]])[["marginal"]]
      after <- model_r2(result$models[[i + 1L]])[["marginal"]]
      path[i, ] %>%
        transmute(
          metric,
          step,
          removed_term = candidate_term,
          R2_before = before,
          R2_after = after,
          delta_R2 = after - before,
          statistic,
          df1,
          df2,
          p.value
        )
    }))
  }))
}

# Print the summary for every metric model.
print_model_summaries <- function(models) {
  for (m in names(models)) {
    metric_header(m)
    print(summary(models[[m]]))
  }
}

# Plot residual diagnostics for each selected model.
plot_model_diagnostics <- function(models) {
  for (metric in names(models)) {
    model <- models[[metric]]
    model_type <- if (inherits(model, "merMod")) "mixed-model" else "OLS"
    metric_header(metric, paste0(" ", model_type, " diagnostics"))
    print(performance::check_model(model))
  }
}

# Apply BH correction and add conventional significance symbols.
add_bh_sig <- function(df, p_col = "p.value") {
  if (!p_col %in% names(df)) {
    stop(sprintf("Column '%s' is required for BH correction.", p_col))
  }
  df %>%
    mutate(
      p_adj_BH = p.adjust(.data[[p_col]], method = "BH"),
      sig = as.character(
        symnum(
          p_adj_BH,
          corr = FALSE,
          na = FALSE,
          cutpoints = c(0, .001, .01, .05, .1, 1),
          symbols = c("***", "**", "*", ".", " ")
        )
      )
    )
}

# Extract marginal and conditional R-squared values where available.
model_r2 <- function(model) {
  if (!inherits(model, "merMod")) {
    return(c(marginal = summary(model)$r.squared, conditional = NA_real_))
  }

  result <- performance::r2(model)
  c(
    marginal = as.numeric(result$R2_marginal[[1L]]),
    conditional = as.numeric(result$R2_conditional[[1L]])
  )
}

# Collect unadjusted contrasts.
collect_simple_effects <- function(models, spec, ...) {
  required_variables <- all.vars(spec)
  eligible_models <- models[vapply(models, function(model) {
    all(required_variables %in% all.vars(formula(model)))
  }, logical(1))]
  if (length(eligible_models) == 0L) {
    return(tibble(
      metric = character(),
      platform = character(),
      contrast = character(),
      estimate = numeric(),
      SE = numeric(),
      lower.CL = numeric(),
      upper.CL = numeric(),
      p.value = numeric()
    ))
  }

  bind_rows(lapply(names(eligible_models), function(m) {
    emmeans(eligible_models[[m]], spec, ...) %>%
      pairs(adjust = "none") %>%
      summary(infer = c(TRUE, TRUE)) %>%
      as.data.frame() %>%
      mutate(metric = m)
  }))
}

# Collect fixed-effect coefficient estimates across metrics.
tidy_fixed_effects <- function(models) {
  bind_rows(lapply(names(models), function(m) {
    model <- models[[m]]
    tidy_result <- if (inherits(model, "merMod")) {
      broom.mixed::tidy(model, effects = "fixed")
    } else {
      broom.mixed::tidy(model)
    }
    tidy_result %>% mutate(metric = m)
  }))
}

# Collect omnibus term tests.
joint_test_table <- function(models) {
  bind_rows(lapply(names(models), function(m) {
    as.data.frame(joint_tests(models[[m]])) %>%
      rename(term = `model term`) %>%
      mutate(metric = m)
  }))
}

# Sort and rejoin the variables in one model term.
sort_model_term <- function(parts) {
  paste(sort(parts), collapse = ":")
}

# Standardise interaction terms so their variable order does not affect matching.
canonical_model_term <- function(term) {
  vapply(
    strsplit(term, ":", fixed = TRUE),
    sort_model_term,
    character(1)
  )
}

# Retain models with a significant BH-adjusted joint test for one term.
models_for_follow_up <- function(models, joint_results, term_name) {
  target_term <- canonical_model_term(term_name)
  eligible_metrics <- joint_results %>%
    filter(
      canonical_model_term(term) == target_term,
      p_adj_BH < 0.05
    ) %>%
    pull(metric)
  models[intersect(names(models), eligible_metrics)]
}

# Print the significance-symbol key.
sig_legend <- function() cat("Signif. codes:  0 '***' 0.001 '**' 0.01 '*' 0.05 '.' 0.1 ' ' 1\n")

# Print every row and column of a result table.
print_all_rows <- function(data) {
  data %>%
    as_tibble() %>%
    print(n = Inf, width = Inf)
}
```

## Backward Elimination -- Stage 1

Simplify random effects first, followed by fixed effects.

```{r stage1-backward-elimination, cache=TRUE}
stage1_backward <- run_backward_selection(STAGE1_FULL_RHS, stage1, METRICS)
stage1_backward_path <- backward_path_table(stage1_backward)
stage1_backward_history <- backward_history_table(stage1_backward)

print_all_rows(stage1_backward_path)
print_all_rows(stage1_backward_history)
print_all_rows(random_elimination_table(stage1_backward))
```

Use the final Stage 1 model for each linguistic measure.

```{r stage1-final-models}
stage1_models <- backward_final_models(stage1_backward)
print(setNames(lapply(stage1_backward, `[[`, "final_rhs"), METRICS))
print(setNames(lapply(stage1_backward, `[[`, "model_type"), METRICS))
```

```{r stage1-results}
print_model_summaries(stage1_models)
```

Check the residual assumptions of each selected model.

```{r stage1-model-diagnostics, fig.width=12, fig.height=12}
plot_model_diagnostics(stage1_models)
```

BH family: all non-intercept coefficients across the four metrics.

```{r stage1-results-BH}
stage1_coefs <- tidy_fixed_effects(stage1_models) %>%
  filter(term != "(Intercept)") %>%
  add_bh_sig("p.value")

bh_header()
print_all_rows(
  stage1_coefs %>% select(metric, term, p.value, p_adj_BH, sig)
)
sig_legend()
```

The following omnibus tests assess complete model terms.

```{r stage1-joint-tests}
stage1_joint <- joint_test_table(stage1_models) %>% add_bh_sig("p.value")

bh_header("omnibus term p-values")
print_all_rows(
  stage1_joint %>%
    select(metric, term, df1, df2, F.ratio, p.value, p_adj_BH, sig)
)
sig_legend()
```

Report platform-specific and equally weighted platform-averaged contrasts.

```{r objective1-gender-within-platform}
gender_by_platform <- collect_simple_effects(
  models_for_follow_up(stage1_models, stage1_joint, "gender:platform"),
  ~ gender | platform,
  weights = "equal"
) %>% add_bh_sig("p.value")

bh_header("gender-within-platform contrast p-values")
print_all_rows(
  gender_by_platform %>%
    select(metric, platform, contrast, estimate, SE, lower.CL, upper.CL, p.value, p_adj_BH, sig)
)
sig_legend()
```

```{r objective1-age-within-platform}
age_by_platform <- collect_simple_effects(
  models_for_follow_up(stage1_models, stage1_joint, "age_group:platform"),
  ~ age_group | platform,
  weights = "equal"
) %>% add_bh_sig("p.value")

bh_header("age_group-within-platform contrast p-values")
print_all_rows(
  age_by_platform %>%
    select(metric, platform, contrast, estimate, SE, lower.CL, upper.CL, p.value, p_adj_BH, sig)
)
sig_legend()
```

```{r objective1-gender-averaged}
gender_averaged <- collect_simple_effects(
  models_for_follow_up(stage1_models, stage1_joint, "gender"),
  ~ gender,
  weights = "equal"
) %>% add_bh_sig("p.value")

bh_header("platform-averaged gender contrast p-values")
print_all_rows(
  gender_averaged %>%
    select(metric, contrast, estimate, SE, lower.CL, upper.CL, p.value, p_adj_BH, sig)
)
sig_legend()
```

```{r objective1-age-averaged}
age_averaged <- collect_simple_effects(
  models_for_follow_up(stage1_models, stage1_joint, "age_group"),
  ~ age_group,
  weights = "equal"
) %>% add_bh_sig("p.value")

bh_header("platform-averaged age_group contrast p-values")
print_all_rows(
  age_averaged %>%
    select(metric, contrast, estimate, SE, lower.CL, upper.CL, p.value, p_adj_BH, sig)
)
sig_legend()
```

## Backward Elimination -- Stage 2

Simplify random effects first, followed by fixed effects.

```{r stage2-backward-elimination, cache=TRUE}
stage2_backward <- run_backward_selection(STAGE2_FULL_RHS, stage2, METRICS)
stage2_backward_path <- backward_path_table(stage2_backward)
stage2_backward_history <- backward_history_table(stage2_backward)

print_all_rows(stage2_backward_path)
print_all_rows(stage2_backward_history)
print_all_rows(random_elimination_table(stage2_backward))
```

Use the final Stage 2 model for each linguistic measure.

```{r stage2-final-models}
stage2_models <- backward_final_models(stage2_backward)
print(setNames(lapply(stage2_backward, `[[`, "final_rhs"), METRICS))
print(setNames(lapply(stage2_backward, `[[`, "model_type"), METRICS))
```

```{r stage2-results}
print_model_summaries(stage2_models)
```

Check the residual assumptions of each selected model.

```{r stage2-model-diagnostics, fig.width=12, fig.height=12}
plot_model_diagnostics(stage2_models)
```

BH family: all non-intercept coefficients across the four metrics.

```{r stage2-results-BH}
stage2_coefs <- tidy_fixed_effects(stage2_models) %>%
  filter(term != "(Intercept)") %>%
  add_bh_sig("p.value")

bh_header()
print_all_rows(
  stage2_coefs %>% select(metric, term, p.value, p_adj_BH, sig)
)
sig_legend()
```

Export the elimination paths, final formulae, and complete candidate histories.

```{r export-backward-selection-results}
SELECTION_OUTPUT_DIR <- "backward_selection_outputs"
dir.create(SELECTION_OUTPUT_DIR, showWarnings = FALSE, recursive = TRUE)

final_formulas <- bind_rows(
  tibble(
    stage = "Stage 1",
    metric = METRICS,
    final_formula = vapply(stage1_backward, `[[`, character(1), "final_rhs")
  ),
  tibble(
    stage = "Stage 2",
    metric = METRICS,
    final_formula = vapply(stage2_backward, `[[`, character(1), "final_rhs")
  )
)

write_csv(stage1_backward_path, file.path(SELECTION_OUTPUT_DIR, "stage1_elimination_path.csv"))
write_csv(stage2_backward_path, file.path(SELECTION_OUTPUT_DIR, "stage2_elimination_path.csv"))
write_csv(stage1_backward_history, file.path(SELECTION_OUTPUT_DIR, "stage1_candidate_history.csv"))
write_csv(stage2_backward_history, file.path(SELECTION_OUTPUT_DIR, "stage2_candidate_history.csv"))
write_csv(final_formulas, file.path(SELECTION_OUTPUT_DIR, "final_formulas.csv"))
write_csv(stage1_coefs, file.path(SELECTION_OUTPUT_DIR, "stage1_final_coefficients.csv"))
write_csv(stage2_coefs, file.path(SELECTION_OUTPUT_DIR, "stage2_final_coefficients.csv"))
```

Omnibus tests for the final selected models:

```{r stage2-joint-tests}
stage2_joint <- joint_test_table(stage2_models) %>% add_bh_sig("p.value")

bh_header("omnibus term p-values")
print_all_rows(
  stage2_joint %>%
    select(metric, term, df1, df2, F.ratio, p.value, p_adj_BH, sig)
)
sig_legend()
```

Equally weighted demographic contrasts:

```{r stage2-gender-averaged}
stage2_gender_averaged <- collect_simple_effects(
  models_for_follow_up(stage2_models, stage2_joint, "gender"),
  ~ gender,
  weights = "equal"
) %>% add_bh_sig("p.value")

bh_header("Stage 2 platform-and-tenure-averaged gender contrast p-values")
print_all_rows(
  stage2_gender_averaged %>%
    select(metric, contrast, estimate, SE, lower.CL, upper.CL, p.value, p_adj_BH, sig)
)
sig_legend()
```

```{r stage2-age-averaged}
stage2_age_averaged <- collect_simple_effects(
  models_for_follow_up(stage2_models, stage2_joint, "age_group"),
  ~ age_group,
  weights = "equal"
) %>% add_bh_sig("p.value")

bh_header("Stage 2 platform-and-tenure-averaged age-group contrast p-values")
print_all_rows(
  stage2_age_averaged %>%
    select(metric, contrast, estimate, SE, lower.CL, upper.CL, p.value, p_adj_BH, sig)
)
sig_legend()
```

Platform-specific demographic contrasts:

```{r stage2-gender-within-platform}
stage2_gender_within_platform <- collect_simple_effects(
  models_for_follow_up(stage2_models, stage2_joint, "gender:platform"),
  ~ gender | platform,
  weights = "equal"
) %>% add_bh_sig("p.value")

bh_header("Stage 2 gender-within-platform contrast p-values")
print_all_rows(
  stage2_gender_within_platform %>%
    select(metric, platform, contrast, estimate, SE, lower.CL, upper.CL, p.value, p_adj_BH, sig)
)
sig_legend()
```

```{r stage2-age-within-platform}
stage2_age_within_platform <- collect_simple_effects(
  models_for_follow_up(stage2_models, stage2_joint, "age_group:platform"),
  ~ age_group | platform,
  weights = "equal"
) %>% add_bh_sig("p.value")

bh_header("Stage 2 age-group-within-platform contrast p-values")
print_all_rows(
  stage2_age_within_platform %>%
    select(metric, platform, contrast, estimate, SE, lower.CL, upper.CL, p.value, p_adj_BH, sig)
)
sig_legend()
```

```{r stage2-age-within-platform-tenure}
stage2_age_within_platform_tenure <- collect_simple_effects(
  models_for_follow_up(
    stage2_models,
    stage2_joint,
    "platform:age_group:tenure"
  ),
  ~ age_group | platform * tenure,
  weights = "equal"
) %>% add_bh_sig("p.value")

bh_header("Stage 2 age-group-within-platform-and-tenure contrast p-values")
print_all_rows(
  stage2_age_within_platform_tenure %>%
    select(
      metric, platform, tenure, contrast, estimate, SE,
      lower.CL, upper.CL, p.value, p_adj_BH, sig
    )
)
sig_legend()
```

Overall and platform-specific tenure contrasts:

```{r stage2-tenure-averaged}
tenure_averaged <- collect_simple_effects(
  models_for_follow_up(stage2_models, stage2_joint, "tenure"),
  ~ tenure,
  weights = "equal"
) %>% add_bh_sig("p.value")

bh_header("platform-averaged tenure contrast p-values")
print_all_rows(
  tenure_averaged %>%
    select(metric, contrast, estimate, SE, lower.CL, upper.CL, p.value, p_adj_BH, sig)
)
sig_legend()
```

```{r stage2-tenure-within-platform}
tenure_within_platform <- collect_simple_effects(
  models_for_follow_up(stage2_models, stage2_joint, "platform:tenure"),
  ~ tenure | platform,
  weights = "equal"
) %>% add_bh_sig("p.value")

bh_header("tenure-within-platform contrast p-values")
print_all_rows(
  tenure_within_platform %>%
    select(metric, platform, contrast, estimate, SE, lower.CL, upper.CL, p.value, p_adj_BH, sig)
)
sig_legend()
```

## Objective 3 — Supporter tenure and demographic variation

The selection histories show whether tenure and its eligible interactions
improved each model. Conditional contrasts are produced only for interactions
retained in a final model.

## Objective 4 — Relative explanatory contributions

Summarise the change associated with each term removed from the Stage 2 models.

```{r objective4-backward-contributions}
stage2_backward_contributions <- backward_contribution_table(stage2_backward)
print_all_rows(stage2_backward_contributions)
```

## Objective 6 — GPT-focused within-between decomposition

```{r objective6-gpt-data}
gpt_data <- df %>% filter(platform == "gpt")
cat(sprintf(
  "GPT-focused EEDI corpus: %d rows (%d stage1 + %d stage2)\n",
  nrow(gpt_data), sum(gpt_data$stage == "stage1"), sum(gpt_data$stage == "stage2")
))

# Standardise metrics before decomposition.
gpt_data <- gpt_data %>%
  mutate(
    z_agency       = as.numeric(scale(agency_ratio)),
    z_formality    = as.numeric(scale(formality_score)),
    z_gainloss     = as.numeric(scale(gain_loss_ratio)),
    z_semdist      = as.numeric(scale(semantic_distance)),
    eedi_score_ord = factor(eedi_score, levels = 1:5, ordered = TRUE)
  )
```

Fit the GPT-focused within--between cumulative link mixed model.

```{r objective6-model, cache=TRUE}
gpt_decomp <- gpt_data %>%
  group_by(condition_id) %>%
  mutate(
    agency_between    = mean(z_agency, na.rm = TRUE),
    agency_within     = z_agency - agency_between,
    formality_between = mean(z_formality, na.rm = TRUE),
    formality_within  = z_formality - formality_between,
    gainloss_between  = mean(z_gainloss, na.rm = TRUE),
    gainloss_within   = z_gainloss - gainloss_between,
    semdist_between   = mean(z_semdist, na.rm = TRUE),
    semdist_within    = z_semdist - semdist_between
  ) %>%
  ungroup()

step1_model <- clmm(
  eedi_score_ord ~ agency_within + agency_between +
    formality_within + formality_between +
    gainloss_within + gainloss_between +
    semdist_within + semdist_between +
    (1 | condition_id),
  data = gpt_decomp, link = "logit", Hess = TRUE
)
summary(step1_model)
```

Check collinearity and approximate the proportional-odds diagnostic using the
equivalent non-mixed model.

```{r objective6-diagnostics}
collinearity_result <- performance::check_collinearity(step1_model)
print(collinearity_result)

step1_clm_check <- clm(
  eedi_score_ord ~ agency_within + agency_between +
    formality_within + formality_between +
    gainloss_within + gainloss_between +
    semdist_within + semdist_between,
  data = gpt_decomp, link = "logit"
)
po_test <- nominal_test(step1_clm_check)
cat("\nProportional-odds check (approximate, via non-mixed clm):\n")
print(po_test)

```
